---
title: "Daily Step Counts"
author: "Karen, Anastasiia, Linda, Asuman, Narod"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(LearnBayes)
source('TeachBayes.r')
library(tidyverse)
```

## Steps per day

We collected a prior from 

We are interested in $\theta$, the average steps per day. 

Prior parameters $\mu_0,\sigma_0$ are calculated based on previous research studies which have been tracked via investigators from the National Cancer Institute (NCI) and the National Institute on Aging (NIA) [daily step counts](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7093766/)

```{r}
# define our prior data set that is collected from previous publications
prior_data_set <- c(3800,4000,3000,4000,8000,11822, 8000,7500,9999,10000,12499)

# plot a histogram
hist(prior_data_set)

# set mu_0 and sigma_0 based on prior data
mu0 <- mean(prior_data_set)
sd0 <- sd(prior_data_set)

# sample size
n <- length(prior_data_set)

```

## Data collection 

We observed daily step counts of a small group of graduate students

a. Store the $k=20$ properties as a vector of average daily step counts in an `R` object called `steps_per_day` 
b. Calculate the mean of these rents and then compute the standard error using $\sigma$ and $n$
c. Use the `dnorm` function to compute the likelihood of observing these data under different values of $\theta$, to see how the likelihood changes 

```{r}
steps_per_day <- c(400, 1000, 1500, 6000, 7016, 7820, 8000, 8000, 8000, 8068, 10000, 10000, 10414, 10953, 11124, 12417, 14000, 14000, 14000, 15000)

k <- length(steps_per_day)
```

