---
title: "Daily Step Counts"
author: "Karen, Anastasiia, Linda, Asuman, Narod"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(LearnBayes)
source('TeachBayes.r')
library(tidyverse)
```

## Prior

We're interested in estimating the average steps of a graduate student per day and how it differs in weekdays and weekend.

Prior parameters $\mu_0,\sigma_0$ are selected based on previous [research studies](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7093766/) which have been tracked via investigators from the National Cancer Institute (NCI) and the National Institute on Aging (NIA).

```{r}
# define our prior data set that is collected from previous publications
# average daily step counts
cohort3_steps <- c(400, 1000, 1500, 6000, 7016, 7820, 8000, 8000, 8000, 8068, 10000, 10000, 10414, 10953, 11124, 12417, 14000, 14000, 14000, 15000, 3800, 4000, 3000, 4000, 8000, 11822,8000, 7500, 9999, 10000, 12499)

# plot a histogram
hist(cohort3_steps, 15)

# set mu_0 and sigma_0 based on prior data
# as we have the average daily step counts, we calculate standard deviation
#mu0 <- mean(prior_data_set)
#sd0 <- sd(prior_data_set)

# sample size
#n <- length(prior_data_set)

```
## Data collection 

We observed average daily step counts of a small group of graduate students

a. Store the $k=21$ properties as a vector of average daily step counts in an `R` object called `steps_per_day` 
b. Calculate the mean of average steps per day and then compute the standard error using $\sigma$ and $k$

```{r}
# daily steps counts of 5G
n1 <- c(11233,16683,19389,1009,8971,9582,5195,9932,16643,7610,7611,15745,7427,7365,6209,15984,17006,11668,11770,9566,14153,8250,14386,18546,19037,11512,9269,5964,8571,8123,10445)
n2 <- c(8632,94,12626,6618,10140,3807,1518,8976,5401,6015,15265,9494,4276,3458,12667,1574,19215,3562,7118,18315,4098,7880,6832,21,10840,12038,13276,13604,9252,13,13473)
n3 <- c(14088,13741,18879,9008,7240,6705,14145,9688,11677,14158,18181,11360,5366,7247,6478,5454,7108,13981,12393,14611,14162,17909,850,658,9519,10773,12942,12354,8212,2122,11823)
n4 <- c(12516,16648,18535,8471,5276,7391,18322,9840,5612,9394,10399,6548,7753,7202,12155,20152,5172,13354,7634,8156,10069,7714,8273,13391,12087,13872,10794,5932,10264,14142,22486)

steps <- c(n1, n2, n3, n4)
students <- c(rep(1,length(n1)),
              rep(2, length(n2)),
              rep(3, length(n3)),
              rep(4, length(n4)))
steps_df <- data.frame(students, steps)
head(steps_df,4)

#plot the daily count steps
ggplot(steps_df, aes(y=steps, group = students)) + 
  geom_boxplot(aes(fill=as.factor(students))) +
  labs(y = "Daily Step Counts" , fill = "Students") + 
  theme_bw()
```

## Hierarchical model in rjags

```{r}
steps_model <- "model{
      # Likelihood model for steps[i]
      for(i in 1:length(steps)) {
          steps[i] ~ dpois(theta[students[i]])
      }

      # Prior models for theta_s
      for(students in 1:n.students) {
      theta[students] ~ dgamma(a, b)
      }

      # Hyperpriors for a,b
      a ~ dgamma(2, 0.2)
      b ~ dgamma(1, 0.5)
  }"

```

## Posterior

```{r}
# ## Calculate shape parameters for beta posterior:
# post_mean <- normal_update(c(mu0,sd0), c(xbar,se))[1]
# post_sd <- normal_update(c(mu0,sd0), c(xbar,se))[2]
# 
# df <- data.frame(Distribution = c("Prior","Data","Posterior"), 
#                  Mean = c(mu0, xbar, post_mean), 
#                  SD = c(sd0, se, post_sd))
# df
```

Plot the prior, data and posterior

```{r}
# x <- seq(min(prior_data_set), max(prior_data_set), by=3000)
# 
# priorx <- dnorm(x, mean=mu0, sd=sd0)
# datax  <- dnorm(x, mean=xbar, sd=se)
# postx  <- dnorm(x, mean=post_mean, sd=post_sd)
# 
# plot(x, priorx, type='l', lwd=3, xlim=c(min(x), max(x)), ylim=c(0,0.0004), col='blue', 
#      main = '', xlab = 'theta', ylab = '')
# lines(x, datax, col='black', lwd=3)
# lines(x, postx, col='red', lwd=3)
# 
# legend("topright", c("Prior","Data","Posterior"), lty = 1, lwd= 3, col = c('blue','black','red'))
```
**The posterior is overlapped with the data**

## Comparing two means

```{r}
# steps_df <- read.csv("steps_df.csv")
# head(steps_df)
```

```{r}
# steps_df %>% group_by(Gender) %>% summarise(group.means = mean(Steps))
```

